SunPy - Python for Solar Physics

### 2014-01-28 3-4-Helioviewer.tex
edits to the Helioviewer setion

SunPy provides the ability to download images hosted by the Helioviewer Project.  
The aim of the Helioviewer Project is to enable the exploration of 
solar and heliospheric data from multiple data sources (such as instrumentation 
and feature/event catalogs) via easy-to-use visual interfaces.  The primary 
data provided by the Helioviewer Project are images in the JPEG2000 format (for 
further details on the Helioviewer Project see \cite{muller2009}). 
SunPy is used in Helioviewer production servers to manage the 
download and ingestion of JPEG2000 files from remote servers.  The JPEG2000 
files are typically highly compressed compared to the source FITS files they 
are generated from, and so can be a used as a shortcut to visualize large 
amounts of data from multiple sources.

The Helioviewer Project categorizes image data based on the physical
construction of the source instrument, using the following hierarchy:
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement, where '$\rightarrow$' means 'provides
(possibly multiple)'.  The full list of image data sources available
can be found through the code snippet, Example
\ref{code:hv:datasources}.
%schriste - what is the point of the following listing? remove?
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> datasources = hv.get_data_sources()

# print a list of datasources and their associated ids
>>> for observatory, instruments in datasources.items():
>>>     for inst, detectors in instruments.items():
>>>         for det, measurements in detectors.items():
>>>             for meas, params in measurements.items():
>>>                 print(observatory, params['nickname'], 
>>>                 params['sourceId']))
\end{minted}
\caption{Initializing the Helioviewer client and listing the full
  range of image data available.}
\label{code:hv:datasources}
\end{listing}

Each Helioviewer Project JPEG2000 file contains metadata which are
based (in part) on the original FITS header information, and carry
sufficient information to permit overlay with other Helioviewer
JPEG2000 files. Images can be accessed either as PNGs (Section
\ref{sssec:hv:png} or as JPEG2000 files (Section \ref{sssec:hv:jp}).

@@

Example \ref{code:hv:downloadlatestpng} demonstrates how to download a
PNG image of the latest AIA 304\AA\ image available on
Helioviewer.org.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

The first argument is the requested time of the image.  Helioviewer
selects images closest to the requested time.  In this case, the
requested time is in the future and so Helioviewer will find the most
recent available image.  The second argument refers to the image
resolution in arcseconds per pixel (larger values mean lower
resolution).  The third argument is a string detailing the requested
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement combination.  The final two numbers in this
string are the visibility and the opacity of the this image layer (1/0
is visible/invisible with opacity in the range $0\rightarrow100$, with
$100$ meaning fully opaque).  The quantities $x0$ and $y0$ are the $x$
and $y$ center points about which to center the image (measured in
helio-projective cartesian co-ordinates) and the \texttt{width} and
\texttt{height} are the pixel values for the image dimensions.

@@

The Helioviewer API allows composition and overlay of images from
multiple sources, based on the positioning metadata in the source FITS
file.  SunPy accesses this overlay/composition capability through the
\texttt{download\_png} method of the Helioviewer client.  Example
\ref{code:hv:overlaid} gives an example of the composition of three
separate images in to a single image.

@@

The principle difference between this code snippet and Example
\ref{code:hv:downloadlatestpng} is in the layer string (the third
argument).  The layer string has three separate comma separated
entries detailing the data source, its visibility and its opacity.
This functionality makes it simple for SunPy users to generate complex
images from multiple, correctly overlaid, image data sources.

@@

Reading JPEG2000 file into a SunPy session requires installing two
other pieces of software. The first, \texttt{OpenJPEG}, is an open
source library for reading and writing JPEG2000 files
(http://www.openjpeg.org).  The other package required is
\texttt{glymur}, (https://github.com/quintusdias/glymur), an interface
between Python and the OpenJPEG libraries (note that these packages
are {\it not} required to use the functionality described in Section
\ref{sssec:hv:png}).

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> filepath = hv.download_jp2('2012/07/05 00:30:00', observatory='SDO',
>>>                            instrument='HMI', detector='HMI', 
>>>                            measurement='continuum')
>>> from sunpy.map import Map
>>> hmi = Map(filepath)
>>> hmi.submap([200,550],[-400,-200]).peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy map object.}
\label{code:downloadjp2}
\end{listing}

>>>>

SunPy provides the ability to download images hosted by the
Helioviewer Project.  The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogs) via
easy-to-use visual interfaces.  The primary data provided by the
Helioviewer Project are images in the JPEG2000 format (for further
details on the Helioviewer Project see \cite{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and so can be a used as a shortcut
to visualize large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

The Helioviewer Project categorizes image data based on the physical
construction of the source instrument, using a simple hierarchy:
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement, where '$\rightarrow$' means 'provides
(possibly multiple)'.  Each Helioviewer Project JPEG2000 file contains
metadata which are based (in part) on the original FITS header
information, and carry sufficient information to permit overlay with
other Helioviewer JPEG2000 files. Images can be accessed either as
PNGs (Section \ref{sssec:hv:png} or as JPEG2000 files (Section
\ref{sssec:hv:jp}).

@@

Listing \ref{code:hv:downloadlatestpng} demonstrates how to download a
PNG image of the latest AIA 304\AA\ image available on
\url{Helioviewer.org}.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

The first argument is the requested time of the image.  Helioviewer
selects images closest to the requested time.  In this case, the
requested time is in the future and so Helioviewer will find the most
recent available image.  The second argument refers to the image
resolution in arcseconds per pixel (larger values mean lower
resolution).  The third argument is a comma-delimited string detailing
the requested observatory $\rightarrow$ instrument $\rightarrow$
detector $\rightarrow$ measurement combination.  The final two numbers
in this string are the visibility and the opacity of the this image
layer (1/0 is visible/invisible with opacity in the range
$0\rightarrow100$, with $100$ meaning fully opaque).  The quantities
$x0$ and $y0$ are the $x$ and $y$ center points about which to center
the image (measured in helio-projective cartesian co-ordinates) and
the \texttt{width} and \texttt{height} are the pixel values for the
image dimensions.

@@

The Helioviewer API allows composition and overlay of images from
multiple sources, based on the positioning metadata in the source FITS
file.  SunPy accesses this overlay/composition capability through the
\texttt{download\_png} method of the Helioviewer client.  Listing
\ref{code:hv:overlaid} gives an example of the composition of three
separate images in to a single image.

@@

The principle difference between this code snippet and Listing
\ref{code:hv:downloadlatestpng} is in the layer string (the third
argument).  The layer string has three separate comma separated
entries detailing the data source, its visibility and its opacity.
This functionality makes it simple for SunPy users to generate complex
images from multiple, correctly overlaid, image data sources.

@@

Reading JPEG2000 file into a SunPy session requires installing two
other pieces of software. The first, \texttt{OpenJPEG}, is an open
source library for reading and writing JPEG2000 files
(\url{http://www.openjpeg.org}).  The other package required is
\texttt{glymur}, (\url{https://github.com/quintusdias/glymur}), an
interface between Python and the OpenJPEG libraries (note that these
packages are {\it not} required to use the functionality described in
Section \ref{sssec:hv:png}).

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> filepath = hv.download_jp2('2012/07/05 00:30:00',
...                                                  observatory='SDO',
...                                                  instrument='HMI', detector='HMI', 
...                                                  measurement='continuum')
>>> from sunpy.map import Map
>>> Map(filepath).submap([200,550],[-400,-200]).peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy map object.}
\label{code:downloadjp2}
\end{listing}
